## GGPLOT

ggplot works with dataframes and not individual vectors

### Understanding basic Syntax

```{r, warning=FALSE, message=FALSE}
## Loading ggplot2
library("ggplot2")

## Removing scientific notation
options(scipen = 999)

data("midwest") ## Load Midwest data

## Init ggplot
ggplot(midwest, aes(x=area, y=poptotal))

```

### How to Make a Simple Scatterplot

```{r, warning=FALSE, message=FALSE}
## Scatter Plot
ggplot(midwest, aes(x=area, y=poptotal)) + geom_point()

## geom_smooth 
g <- ggplot(midwest, aes(x=area, y=poptotal)) + geom_point() + geom_smooth(method="lm")
plot(g)

```

### Adjusting the X and Y axis limits

#### Method 1 - By deleting the points outside the range

```{r, warning=FALSE, message=FALSE}
g + xlim(c(0, 0.1)) + ylim(c(0, 1000000))

```

Did you notice that the line of best fit became more horizontal compared to the original
plot? This is because, when using xlim() and ylim(), the points outside the specified
range are deleted and will not be considered while drawing the line of best fit (using
geom_smooth(method='lm')).

#### Method 2 - Zooming In

```{r, warning=FALSE, message=FALSE}
g1 <- g + coord_cartesian(xlim=c(0,0.1), ylim=c(0, 1000000))
plot(g1)

```

Since all points were considered, the line of best fit did not change

### How to Change the Title and Axis Labels

```{r, warning=FALSE, message=FALSE}
g1 + labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population", x="Area", caption="Midwest Demographics")

```

### How to Change the Color and Size of Points

#### How to Change the Color and Size To Static

```{r, warning=FALSE, message=FALSE}

ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(col="steelblue", size=3) +   # Set static color and size for points
  geom_smooth(method="lm", col="firebrick") +  # change the color of line
  coord_cartesian(xlim=c(0, 0.1), ylim=c(0, 1000000)) + 
  labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population", x="Area", caption="Midwest Demographics")

```

#### How to Change the Color To Reflect Categories in Another Column

```{r, warning=FALSE, message=FALSE}
gg <- ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state), size=3) +  # Set color to vary based on state categories.
  geom_smooth(method="lm", col="firebrick", size=2) + 
  coord_cartesian(xlim=c(0, 0.1), ylim=c(0, 1000000)) + 
  labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population", x="Area", caption="Midwest Demographics")
plot(gg)

```


### How to Change the X Axis Texts and Ticks Location

#### How to Change the X and Y Axis Text and its Location

Step 1: Set the breaks

The breaks should be of the same scale as the X axis variable. Note that I am using
scale_x_continuous because, the X axis variable is a continuous variable. Had it been a
date variable, scale_x_date could be used

```{r, warning=FALSE, message=FALSE}

gg + scale_x_continuous(breaks=seq(0, 0.1, 0.01))

```

Step 2: Change the labels

```{r, warning=FALSE, message=FALSE}

gg + scale_x_continuous(breaks=seq(0, 0.1, 0.01), labels = letters[1:11])

```

#### How to Write Customized Texts for Axis Labels, by Formatting the Original Values

```{r, warning=FALSE, message=FALSE}

gg + scale_x_continuous(breaks=seq(0, 0.1, 0.01), labels = sprintf("%1.2f%%", seq(0, 0.1, 0.01))) + 
  scale_y_continuous(breaks=seq(0, 1000000, 200000), labels = function(x){paste0(x/1000, 'K')})

```
